Neighbor-Based Decentralized Training Strategies for Multi-Agent Reinforcement Learning

### 2024-10-02 paper-2024-sac-marl-neighbouring-strategies.tex
adding something, i dunno what, i'm really tired and I want to rasengan
more in sota

\paragraph{Centralized Training with Decentralized Execution (CTDE)}
In CTDE, a central learner trains all agents, 
but execution of the learned policy is decentralized. 
This process typically involves two phases:
\begin{itemize}
    \item \textbf{Offline training:} agents interact with a simulator or environment, 
    gathering experience data used to update the central learner's policy.
    \item \textbf{Online execution:} agents deploy the learned policy in the environment without further learning.
\end{itemize}
This approach offers several benefits.  
Leveraging a centralized dataset allows for efficient learning, 
potentially exploiting global information, 
which is crucial in cooperative scenarios requiring coordinated actions. 
Furthermore, decentralized execution allows agents to act independently after training.
%
However, CTDE has limitations. 
Scalability can become problematic with a large number of agents. 
The reliance on a central learner hinders adaptation to dynamic environments or tasks during runtime.
\sloppy
\paragraph{Decentralized Training with Decentralized Execution (DTDE)}
DTDE eliminates the central learner; each agent learns independently. 
Inter-agent cooperation and communication can be incorporated during both training and execution. 
Unlike CTDE, training and execution phases are not necessarily separate. 
Simultaneous learning and execution allow for adaptation to changing environments or tasks at runtime.
%
Key advantages of DTDE include scalability with the number of agents and adaptability to dynamic environments.
%
However, DTDE faces challenges. 
Convergence time is often longer compared to centralized training. 
The lack of central coordination can lead to instability and difficulties in differentiating agents and their environmental interactions. 
Online learning in DTDE can introduce non-stationarity if agents do not account for other agents' evolving policies during training, 
as the environment effectively changes from each agent's perspective.
\subsection{State of the Art}
Multi-agent reinforcement learning is a longstanding research area with a rich literature, starting 
from seminal works like [cite ] and [cite] to more recent contributions like [cite] and [cite].
%
In the following, we briefly discuss some of the most relevant approaches in the field.
\paragraph{Independent Learners}

@@

In [ref] we provide a detailed description of the integration of the neighboring-based strategies with deep Q-learning.
\begin{algorithm}
  \caption{Neighboring-Based Deep Q-Learning}
  \label{alg:neighbor_dqn}
  \ForEach{agent $i \in N$}{
      Initialize Q-network $Q^i(o^i, a^i; \theta^i)$ and target network $Q^{i-}(o^i, a^i; \theta^{i-})$\\
      Initialize replay buffer $\mathcal{B}^i$
  }
  \For{episode = 1 to M}{
      \ForEach{agent $i \in N$}{
          Initialize initial observation $o_0^i$
      }
      \For{t = 0 to T}{
          \ForEach{agent $i \in N$}{
              Select action $a_t^i$ based on $\epsilon$-greedy policy using $Q^i(o_t^i, a^i; \theta^i)$\\
              Execute action $a_t^i$ and observe reward $r_{t+1}^i$ and next observation $o_{t+1}^i$\\
              Store $(o_t^i, a_t^i, r_{t+1}^i, o_{t+1}^i)$ in $\mathcal{B}^i$\\
              Sample a mini-batch from $\mathcal{B}^i$\\
              Calculate target values using chosen neighboring strategy (see below)\\
              Update $\theta^i$ by minimizing the TD error\\
              Every C steps, update target network $\theta^{i-} \leftarrow \theta^i$\\
              \textbf{
              Every S steps, perform neighbor interaction (e.g., averaging, consensus, experience sharing)}
          }
      }
  }
\end{algorithm}

>>>>

\paragraph{Centralized Training with Decentralized Execution (CTDE)}
In CTDE, a central learner trains all agents, 
but execution of the learned policy is decentralized. 
This process typically involves two phases:
\begin{itemize}
    \item \textbf{Offline training:} agents interact with a simulator or environment, 
    gathering experience data used to update the central learner's policy.
    \item \textbf{Online execution:} agents deploy the learned policy in the environment without further learning.
\end{itemize}
This approach offers several benefits.  
Leveraging a centralized dataset allows for efficient learning, 
potentially exploiting global information, 
which is crucial in cooperative scenarios requiring coordinated actions. 
Furthermore, decentralized execution allows agents to act independently after training.
%
However, CTDE has limitations. 
Scalability can become problematic with a large number of agents. 
The reliance on a central learner hinders adaptation to dynamic environments or tasks during runtime.
\sloppy
\paragraph{Decentralized Training with Decentralized Execution (DTDE)}
DTDE eliminates the central learner; each agent learns independently. 
Inter-agent cooperation and communication can be incorporated during both training and execution. 
Unlike CTDE, training and execution phases are not necessarily separate. 
Simultaneous learning and execution allow for adaptation to changing environments or tasks at runtime.
%
Key advantages of DTDE include scalability with the number of agents and adaptability to dynamic environments.
%
However, DTDE faces challenges. 
Convergence time is often longer compared to centralized training. 
The lack of central coordination can lead to instability and difficulties in differentiating agents and their environmental interactions. 
Online learning in DTDE can introduce non-stationarity if agents do not account for other agents' evolving policies during training, 
as the environment effectively changes from each agent's perspective.
\subsection{State of the Art}
Multi-agent reinforcement learning is a longstanding research area with a rich literature, starting 
from seminal works like~\cite{tan1993multi} and~\cite{busoniu2008comprehensive} to more recent contributions like~\cite{gronauer2022multi} and~\cite{canese2021multi}.
%
In the following, we briefly discuss some of the most relevant approaches in the field.
\paragraph{Independent Learners}

@@

In \Cref{alg:neighbor_dqn} we provide a detailed description of the integration of the neighboring-based strategies with deep Q-learning.
\begin{algorithm}
  \caption{Neighboring-Based Deep Q-Learning}
  \label{alg:neighbor_dqn}
  \ForEach{agent $i \in N$}{
      Initialize Q-network $Q^i(o^i, a^i; \theta^i)$ and target network $Q^{i-}(o^i, a^i; \theta^{i-})$\\
      Initialize replay buffer $\mathcal{B}^i$
  }
  \For{episode = 1 to M}{
      \ForEach{agent $i \in N$}{
          Initialize initial observation $o_0^i$
      }
      \For{t = 0 to T}{
          \ForEach{agent $i \in N$}{
              Select action $a_t^i$ based on $\epsilon$-greedy policy using $Q^i(o_t^i, a^i; \theta^i)$\\
              Execute action $a_t^i$ and observe reward $r_{t+1}^i$ and next observation $o_{t+1}^i$\\
              Store $(o_t^i, a_t^i, r_{t+1}^i, o_{t+1}^i)$ in $\mathcal{B}^i$\\
              Sample a mini-batch from $\mathcal{B}^i$\\
              Calculate target values using chosen neighboring strategy (see below)\\
              Update $\theta^i$ by minimizing the TD error\\
              Every C steps, update target network $\theta^{i-} \leftarrow \theta^i$\\
              \textbf{
              Every S steps, perform neighbor interaction (e.g., averaging, consensus, experience sharing)}
          }
      }
  }
\end{algorithm}
